Hippocampal segmentation with MAGeT

### 2012-11-14 paper.Rnw
Added the SZ dataset results -- MAGeT vs. Manual volumes.

\begin{document}
\SweaveOpts{concordance=TRUE}

\title{Hippocampal segmentation with MAGeT}
\author{Jon Pipitone, }
\maketitle

@@

\begin{figure}[h]
  \centering
    \includegraphics[width=0.5\textwidth]{MAGeT-figure.png}
  \caption{Diagram of the MAGeT Brain algorithm}
\end{figure}

@@

<<a2a, cache=TRUE, fig = T, echo = F, dev='CairoPDF', warning=F, error=F>>=
library(ggplot2)
multiplot <- function(..., plotlist=NULL, cols) {
    require(grid)

@@

#loess.control(trace.hat = c("approximate"), statistics = c("approximate"), surface = c("interpolate"))
pdf("a2a_kbytemplates.pdf", width=15, height=15)
multiplot(t_maj_plot  , a_maj_plot  ,
          t_xcorr_plot, a_xcorr_plot, 
          t_nmi_plot  , a_nmi_plot  , cols=2)
dev.off()

@@

- more atlases -> better performance
- larger template library -> better performance, but tails off around 10-15 templates
- no significant difference between majority or weighted vote methods (haven’t tested this statistically though). 
- consistently performs better than average naive performance by XXX
- using ANTS, with a large enough template library (>12) MAGeT brain performs better than the average multi-atlas approach with the same number of atlases.  using ANIMAL, 5 or more atlases needed before boost seen. 
- more atlases -> smaller template library required to improve on average multi-atlas performance
- discuss variance?  best/worst case? --how often do we expect random template library selection to work decently

@@

- A2A shows that if atlas population strongly(?) represents subject set variability, then free choice from atlas population will produce improvements (we know this b/c of extensive validation trials). 
- what about in the case where atlas population doesn’t strongly represent subject set variability (e.g. a priori atlas set)?  then, we can use atlas selection to refine atlas set?

@@

\begin{figure}[h]
<<ADNI-baseline-volumes-boxplot,echo=F,fig=T, fig.align='center',dependson='ADNI-baseline-volumes-prep',cache=T>>=
library(reshape2)
qplot(DX,value,data=melt(totals,id.vars=c("ID","DX"),measure.vars=c("FSL","MAPER","MAGeT","SNT"),na.rm=T),colour=variable,geom="boxplot")
@
  \caption{Comparison of HC volumes by FreeSurfer (FSF), MAGeT brain (MAGeT), MAPER, and manual (SNT).}
  \label{ADNI-baseline-volumes-boxplot}
\end{figure}

\begin{figure}[h]
<<ADNI-baseline-volumes-plot,echo=F,fig=T, fig.align='center',dependson='ADNI-baseline-volumes',cache=T>>=
library(reshape2)
qplot(SNT,value,data=melt(totals,id.vars=c("ID","SNT"),measure.vars=c("FSL","MAPER","MAGeT")),colour=variable, geom="smooth", method="lm")
@
  \caption{Comparison of HC volumes by FreeSurfer (FSF), MAGeT brain (MAGeT), MAPER, and manual (SNT).}
  \label{ADNI-baseline-volumes-plot}
\end{figure}

\end{document}

>>>>

\begin{document}
\SweaveOpts{echo = TRUE, message = FALSE, warning = FALSE, fig.align = 'center', tidy = FALSE, comment = NA, cache = TRUE}

<<loadpackages, results="hide", echo=TRUE, cache=FALSE>>=
library(ggplot2)
library(Hmisc)
library(reshape2)
@

\title{Hippocampal segmentation with MAGeT}
\author{Jon Pipitone}
\maketitle

@@

\begin{figure}[h]
  \centering
    \includegraphics[width=0.5\textwidth]{figure/MAGeT-figure.png}
  \caption{Diagram of the MAGeT Brain algorithm}
\end{figure}

@@

<<a2a, cache=TRUE, fig = T, echo = F, dev='CairoPDF', warning=F, error=F>>=
multiplot <- function(..., plotlist=NULL, cols) {
    require(grid)

@@

#loess.control(trace.hat = c("approximate"), statistics = c("approximate"), surface = c("interpolate"))
#pdf("a2a_kbytemplates.pdf", width=15, height=15)
multiplot(t_maj_plot  , a_maj_plot  ,
          t_xcorr_plot, a_xcorr_plot, 
          t_nmi_plot  , a_nmi_plot  , cols=2)
#dev.off()

@@

- more atlases -> better performance
- larger template library -> better performance, but tails off around 10-15 templates
- no significant difference between majority or weighted vote methods (haven???t tested this statistically though). 
- consistently performs better than average naive performance by XXX
- using ANTS, with a large enough template library (>12) MAGeT brain performs better than the average multi-atlas approach with the same number of atlases.  using ANIMAL, 5 or more atlases needed before boost seen. 
- more atlases -> smaller template library required to improve on average multi-atlas performance
- discuss variance?  best/worst case? --how often do we expect random template library selection to work decently

@@

- A2A shows that if atlas population strongly(?) represents subject set variability, then free choice from atlas population will produce improvements (we know this b/c of extensive validation trials). 
- what about in the case where atlas population doesn???t strongly represent subject set variability (e.g. a priori atlas set)?  then, we can use atlas selection to refine atlas set?

@@

\begin{figure}[h]
<<ADNI-baseline-volumes-boxplot,echo=F,fig=T, fig.align='center',dependson='ADNI-baseline-volumes-prep',cache=T>>=
qplot(DX,value,data=melt(totals,id.vars=c("ID","DX"),measure.vars=c("FSL","MAPER","MAGeT","SNT"),na.rm=T),colour=variable,geom="boxplot")
@
  \caption{Comparison of HC volumes by FreeSurfer (FSF), MAGeT brain (MAGeT), MAPER, and manual (SNT).}
  \label{ADNI-baseline-volumes-boxplot}
\end{figure}

\begin{figure}[h]
<<ADNI-baseline-volumes-plot,echo=F,fig=T, fig.align='center',dependson='ADNI-baseline-volumes',cache=T>>=
qplot(SNT,value,data=melt(totals,id.vars=c("ID","SNT"),measure.vars=c("FSL","MAPER","MAGeT")),colour=variable, geom="smooth", method="lm")
@
  \caption{Comparison of HC volumes by FreeSurfer (FSF), MAGeT brain (MAGeT), MAPER, and manual (SNT).}
  \label{ADNI-baseline-volumes-plot}
\end{figure}

\subsubsection{First Episode Schizophrenic patients}

<<FEP-SZ-demographics, echo=F, results="asis",cache=FALSE>>=
fep_sz_demographics = read.csv("data/SZ_demographics.csv", na.strings=c("","."))
fep_sz_demographics = subset(fep_sz_demographics, Anon <= 81)
tab <- summary(DX ~ Age + Gender + Handedness + Education + SES + FSIQ, data=fep_sz_demographics, method = "reverse", test = FALSE)
latex(tab, file="", size="scriptsize", where="h")
@

\begin{figure}[h]
<<SZ_volumes,echo=F,fig=T, fig.align='center',dependson='ADNI-baseline-volumes',cache=T>>=
SZ_volumes=read.csv("data/SZ_volumes_by_method.csv")
qplot(Manual,MAGeT,data=SZ_volumes) + geom_smooth(method="lm")
@
  \caption{Comparison of total HC volumes for MAGeT against manually rated volumes}
  \label{SZ_volumes}
\end{figure}
\end{document}

### 2012-11-14 paper.tex
Added the SZ dataset results -- MAGeT vs. Manual volumes.

\title{Hippocampal segmentation with MAGeT}
\author{Jon Pipitone, }
\maketitle

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}

{\ttfamily\noindent\itshape\textcolor{messagecolor}{\#\# geom\_smooth: method="auto" and size of largest group is >=1000, so using gam with formula: y \textasciitilde{} s(x, bs = "cs"). Use 'method = x' to change the smoothing method.}}

{\ttfamily\noindent\itshape\textcolor{messagecolor}{\#\# geom\_smooth: method="auto" and size of largest group is >=1000, so using gam with formula: y \textasciitilde{} s(x, bs = "cs"). Use 'method = x' to change the smoothing method.}}

{\ttfamily\noindent\itshape\textcolor{messagecolor}{\#\# geom\_smooth: method="auto" and size of largest group is >=1000, so using gam with formula: y \textasciitilde{} s(x, bs = "cs"). Use 'method = x' to change the smoothing method.}}

{\ttfamily\noindent\itshape\textcolor{messagecolor}{\#\# geom\_smooth: method="auto" and size of largest group is >=1000, so using gam with formula: y \textasciitilde{} s(x, bs = "cs"). Use 'method = x' to change the smoothing method.}}

{\ttfamily\noindent\itshape\textcolor{messagecolor}{\#\# geom\_smooth: method="auto" and size of largest group is >=1000, so using gam with formula: y \textasciitilde{} s(x, bs = "cs"). Use 'method = x' to change the smoothing method.}}

{\ttfamily\noindent\itshape\textcolor{messagecolor}{\#\# geom\_smooth: method="auto" and size of largest group is >=1000, so using gam with formula: y \textasciitilde{} s(x, bs = "cs"). Use 'method = x' to change the smoothing method.}}\begin{verbatim}
## pdf 
##   2
\end{verbatim}

{\ttfamily\noindent\itshape\textcolor{messagecolor}{\#\# geom\_smooth: method="auto" and size of largest group is >=1000, so using gam with formula: y \textasciitilde{} s(x, bs = "cs"). Use 'method = x' to change the smoothing method.}}

{\ttfamily\noindent\itshape\textcolor{messagecolor}{\#\# geom\_smooth: method="auto" and size of largest group is >=1000, so using gam with formula: y \textasciitilde{} s(x, bs = "cs"). Use 'method = x' to change the smoothing method.}}

{\ttfamily\noindent\itshape\textcolor{messagecolor}{\#\# geom\_smooth: method="auto" and size of largest group is >=1000, so using gam with formula: y \textasciitilde{} s(x, bs = "cs"). Use 'method = x' to change the smoothing method.}}

{\ttfamily\noindent\itshape\textcolor{messagecolor}{\#\# geom\_smooth: method="auto" and size of largest group is >=1000, so using gam with formula: y \textasciitilde{} s(x, bs = "cs"). Use 'method = x' to change the smoothing method.}}

{\ttfamily\noindent\itshape\textcolor{messagecolor}{\#\# geom\_smooth: method="auto" and size of largest group is >=1000, so using gam with formula: y \textasciitilde{} s(x, bs = "cs"). Use 'method = x' to change the smoothing method.}}

{\ttfamily\noindent\itshape\textcolor{messagecolor}{\#\# geom\_smooth: method="auto" and size of largest group is >=1000, so using gam with formula: y \textasciitilde{} s(x, bs = "cs"). Use 'method = x' to change the smoothing method.}}\end{kframe}\includegraphics[width=\maxwidth]{figure/a2a} 
\end{knitrout}

@@

\begin{figure}[h]
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}

{\ttfamily\noindent\textcolor{warningcolor}{\#\# Warning: Removed 359 rows containing missing values (stat\_smooth).}}

{\ttfamily\noindent\textcolor{warningcolor}{\#\# Warning: Removed 355 rows containing missing values (stat\_smooth).}}

{\ttfamily\noindent\textcolor{warningcolor}{\#\# Warning: Removed 355 rows containing missing values (stat\_smooth).}}\end{kframe}

@@

\end{document}

>>>>

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlfunctioncall{library}(ggplot2)
\hlfunctioncall{library}(Hmisc)
\hlfunctioncall{library}(reshape2)
\end{alltt}
\end{kframe}
\end{knitrout}

\title{Hippocampal segmentation with MAGeT}
\author{Jon Pipitone}
\maketitle

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}

{\centering \includegraphics[width=\maxwidth]{figure/a2a1}

{\centering \includegraphics[width=\maxwidth]{figure/a2a2}

@@

@@

\subsubsection{First Episode Schizophrenic patients}

% latex.default(cstats, title = title, caption = caption, rowlabel = rowlabel,      col.just = col.just, numeric.dollar = FALSE, insert.bottom = legend,      rowname = lab, dcolumn = dcolumn, extracolheads = extracolheads,      extracolsize = Nsize, ...) 
%
\begin{table}[h]
\scriptsize
\caption{Descriptive Statistics by DX\label{tab}} 
\begin{center}
\begin{tabular}{lrc}
\hline\hline
\multicolumn{1}{l}{}&\multicolumn{1}{c}{N}&\multicolumn{1}{c}{FEP}\tabularnewline
&&\multicolumn{1}{c}{{\scriptsize $N=81$}}\tabularnewline
\hline
Age&80&{\scriptsize 21~}{23 }{\scriptsize 26} \tabularnewline
Gender~:~M&81&63\%~{\scriptsize~(51)}\tabularnewline
Handedness~:~ambi&81&~6\%~{\scriptsize~(~5)}\tabularnewline
~~~~left&&~5\%~{\scriptsize~(~4)}\tabularnewline
~~~~right&&89\%~{\scriptsize~(72)}\tabularnewline
Education&81&{\scriptsize 11~}{13 }{\scriptsize 15} \tabularnewline
SES~:~lower&81&31\%~{\scriptsize~(25)}\tabularnewline
~~~~middle&&54\%~{\scriptsize~(44)}\tabularnewline
~~~~upper&&15\%~{\scriptsize~(12)}\tabularnewline
FSIQ&79&{\scriptsize  88~}{102 }{\scriptsize 109} \tabularnewline
\hline
\end{tabular}
\end{center}
\noindent {\scriptsize $a$\ }{$b$\ }{\scriptsize $c$\ } represent the lower quartile $a$, the median $b$, and the upper quartile $c$\ for continuous variables.\\$N$\ is the number of non--missing values.\\Numbers after percents are frequencies.\end{table}

{\centering \includegraphics[width=\maxwidth]{figure/SZ_volumes}

  \caption{Comparison of total HC volumes for MAGeT against manually rated volumes}
  \label{SZ_volumes}
\end{figure}
\end{document}
